API design for machine learning software: experiences from the scikit-learn project

### 2013-10-18 paper.tex
Move snippets again.
Copy edit implementation and related work.

Our implementation guidelines emphasize writing efficient but readable code.
In particular, we focus on making the codebase maintainable and
understandable in order to favor external contributions. Whenever practical,
algorithms implemented in \sklearn are written in Python,
using NumPy vector operations for numerical work.
This allows the code to remain concise, readable and efficient.
For critical algorithms that cannot be easily and efficiently
expressed as NumPy operations, we rely on Cython \citep{behnel2011cython}
to achieve competitive performance and scalability. Cython is a
compiled programming language that extends Python with static typing. It
produces efficient C extension modules that are importable from the
Python run-time system. Examples of Cython code in \sklearn are
stochastic gradient descent for linear models, some clustering
algorithms, and decision trees.

@@

\sklearn competes with a many machine learning packages,
both open source and proprietary.
Some of those, including Weka~\citep{hall2009weka} or
Orange~\citep{Demsar2004}, offer APIs but actually focus on the use of a graphical user interface (GUI)
which allows novices to easily apply machine learning algorithms. By
contrast, the target audience of \sklearn is capable of programming, and
therefore we focus on developing a usable and consistent API, rather than expend
effort into creating a GUI\@. In addition, while GUIs are useful tools, they
sometimes make reproducibility difficult in the case of complex workflows
(although those packages usually have developed a GUI for managing complex
tasks).

@@

Within the realm of Python, two packages deserve mention.
The Pylearn2 library \citep{goodfellow2013pylearn2}
is similar to \sklearn in that aims to facilitate reproducible results
and model inspection.
However, it is designed for the audience of machine learning researchers,
rather than practitioners in other fields.
This leads to a modular toolkit,
where each element of an algorithm can be tweaked or replaced as needed.
This provides more control for the user,
at the cost of more math to be mastered.

Finally, the Gensim topic modeling toolkit \citep{rehurek2010gensim}
exemplifies an API geared toward ``big data'' processing.
Gensim's method of dealing with large datasets is to use algorithms
that have small space complexity and can be updated online.
The API is designed around the Python concept of an ``iterable''
(supported in the language by a restricted form of co-routines called generators).
Some \sklearn estimators, notably Naïve Bayes and Stochastic Gradient Descent
(for learning linear models such as
logistic regression and SVM),
have a \texttt{partial\_fit} method to learn from a stream of samples
that is divided into mini-batches (chunks).
This way of processing, as argued in Section~\ref{sec:arrays},
allows for more efficient implementation.
We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory dedicated to \sklearn algorithms.

>>>>

Our implementation guidelines emphasize writing efficient but readable code.
In particular, we focus on making the codebase maintainable and
understandable in order to favor external contributions. Whenever practical,
algorithms implemented in \sklearn are written in Python,
using NumPy vector operations for numerical work.
This allows the code to remain concise, readable and efficient.
For critical algorithms that cannot be easily and efficiently
expressed as NumPy operations, we rely on Cython \citep{behnel2011cython}
to achieve competitive performance and scalability. Cython is a
compiled programming language that extends Python with static typing. It
produces efficient C extension modules that are importable from the
Python run-time system. Examples of Cython code in \sklearn are
Stochastic Gradient Descent for linear models, some clustering
algorithms, and decision trees.

@@

\sklearn competes with many machine learning packages,
both open source and proprietary.
Some of those, including Weka~\citep{hall2009weka} or
Orange~\citep{Demsar2004}, offer APIs but actually focus on the use of a graphical user interface (GUI)
which allows novices to easily apply machine learning algorithms. By
contrast, the target audience of \sklearn is capable of programming, and
therefore we focus on developing a usable and consistent API, rather than expend
effort into creating a GUI\@. In addition, while GUIs are useful tools, they
sometimes make reproducibility difficult in the case of complex workflows
(although those packages usually have developed a GUI for managing complex
tasks).

@@

Within the realm of Python, two packages deserve mention.
The Pylearn2 library \citep{goodfellow2013pylearn2}
is similar to \sklearn in that it aims to facilitate reproducible results
and model inspection.
However, it is designed for the audience of machine learning researchers,
rather than practitioners in other fields.
This leads to a modular toolkit,
where each element of an algorithm can be tweaked or replaced as needed.
This provides more control for the user,
at the cost of more advanced concepts to be mastered.

Finally, the Gensim topic modeling toolkit \citep{rehurek2010gensim}
exemplifies an API geared toward ``big data'' processing.
Gensim's method of dealing with large datasets is to use algorithms
that have small space complexity and can be updated online.
The API is designed around the Python concept of an ``iterable''
(supported in the language by a restricted form of co-routines called generators).
Some \sklearn estimators, notably Naïve Bayes and Stochastic Gradient Descent,
have a \texttt{partial\_fit} method to learn from a stream of samples
that is divided into mini-batches (chunks).
This way of processing, as argued in Section~\ref{sec:arrays},
allows for more efficient implementation.
We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory dedicated to \sklearn algorithms.
